Renewal and stability in populations structured by remaining years of life

### 2015-11-10 ThanoRepro/ThanoRepro.tex
version submitted to DR

\documentclass{article}
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{pgf}
\usepackage{tikz}
%\usetikzlibrary{arrows}
%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\usepackage{url}
\usepackage{nth}
% for the d in integrals
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\ec}{\quad\quad\text{,}}
\newcommand{\ep}{\quad\quad\text{.}}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{setspace}
\newcommand\ackn[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

@@

\title{Renewal and stability in populations structured by remaining years of
life}
\author{Tim Riffe \\ Max Planck Institute for Demographic Research}
\maketitle

\begin{abstract}
Models of age structured population renewal are the core of formal demography.
I derive a specification of the classic renewal equation that is structured by
thanatological age rather than by chronological age. I give both continuous
and discrete variants of the derived model, and relate these to the well-known
age structured model. In stability, the two kinds of age
structure are commensurable, and imply identical intrinsic growth rates.
I demonstrate approximate symmetry between chronological and
thanatological age structure in stability when subject to
intrinsic growth rates equal magnitude and opposite sign. This paper offers a
new perspective on the Lotka-Leslie renewal process.\ackn{Research reported in
this manuscript was supported by the U.S.
National Institute On Aging of the National Institutes of Health under award numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the
author and does not necessarily represent the official views of the funding
agencies.}
\end{abstract}

\onehalfspacing
\section*{Introduction}

Demography is built upon a small set of empirical
regularities with respect to chronological age patterns. The models and methods
of demography are in varying degrees prefaced on such regularity. There is evidence that many demographic phenomena are best described in the
aggregatate as a function of time since birth, and others of time until death,
while others can be a function of both to some degree
\citep{riffe2015ttd,wolf2015disability}. This observation motivates
incorporating this aspect time-to-death into formal demographic methods and
models.

>>>>

\documentclass[12pt,oneside,a4paper,leqno]{article}
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{placeins}
\usepackage{graphicx}
\usepackage{subcaption}
%\usepackage{pgf}
\usepackage{tikz}

%\setlength\PreviewBorder{5pt}
\usepackage{natbib}
\bibpunct{(}{)}{,}{a}{}{;} 
\usepackage{url}
\usepackage{nth}
% for the d in integrals
\newcommand{\dd}{\; \mathrm{d}}
\newcommand{\ec}{\quad\quad\text{,}}
\newcommand{\ep}{\quad\quad\text{.}}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{setspace}
\doublespacing

\newcommand\ackn[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

@@

\title{Renewal and stability in populations structured by remaining years of
life}
\author{(author redacted)}
\maketitle

%\ackn{Research
%reported in this manuscript was supported by the U.S.
%National Institute On Aging of the National Institutes of Health under award
% numbers R01-AG011552 and R01-AG040245. The content is solely the responsibility of the author and does not necessarily represent the official views of the funding
%agency.}

\section*{Structured abstract}

\subsection*{Background}
The Lotka-Leslie renewal model is the core of formal demography. This model is
structured by chronological age, and it does not account for thanatological age.

\subsection*{Objective}
I derive a specification of the classic renewal equation that is structured by
thanatological age rather than by chronological age. I give both continuous
and discrete variants of the derived model, and relate these to the Lotka-Leslie
renewal model.

\subsection*{Results}
In stability, the thanatological and chronological renewal models are
commensurable, implying identical intrinsic growth rates.
I demonstrate approximate symmetry between chronological and
thanatological age structure in stability when subject to
intrinsic growth rates equal magnitude and opposite sign.

\subsection*{Conclusions}
Birth-death renewal processes can be expressed as death-birth processes, and
vice versa.

\subsection*{Contribution}
The thanatological renewal model offers a new perspective on population renewal,
and it is valid more generally as an aspect of birth-death processes.
\vspace{2cm}

\section*{Introduction}

Contemporary demography is built upon a small set of empirical
regularities with respect to chronological age patterns. The models and methods
of demography are in varying degrees prefaced on such regularity. There is evidence that many demographic phenomena are best described in the
aggregatate as a function of time since birth, and others of time until death,
while others can be a function of both to some degree
\citep{riffe2015ttd,wolf2015disability}. This observation motivates
incorporating this aspect time-to-death into formal demographic methods and
models.
